---
title: "Education Data Science Summit"
author: "Ryan Estrellado"
date: "4/20/2021"
output: html_document
---

# What is an R Markdown Document? 

The authors of **R Markdown: The Definitive Guide** write that R Markdown empowers data scientists to "interweave narratives with code in a document . . ." 

Think of it as a Word document that holds not only story, but also code and graphs. 


# Things to Watch For 

Look for how writing, code, and plots appear in a single document. 

Look for the parts of good dat analysis using code: 

 - Identify the Question 
 - Import the Data 
 - Prepare the Data 
 - Pick a method to answer the question 
 - Visualize the data 
 - Analyze the Data 
 - Report the findings 


# Load Packages 

```{r load packages}
library(tidyverse)
library(here)
library(janitor)
```


# Identify the Question 

How many schools do districts have on average? 


# Import Data 

Before you work with data in R, you have to import the data. 

You can get this data from the [California Department of Education's website](https://www.cde.ca.gov/ds/sd/sd/filesenr.asp): 

```{r import data}
enroll <- read_tsv(here::here("19-20-enrollment.txt"))
```

Here's one way working in R is different from working in a spreadsheet: you aren't dragging and dropping things around in a spreadsheet. Instead, you're using code to submit instructions to R. Sometimes you just want to see the dataset. Here's a way to do that: 

```{r view data}
# See the data using View()
View(enroll)
```

Question: Let's look at a few variables. Do these variables contain unique values? In other words, does Castle Rock school appear only once? 


# Prepare Data 

Most publicly available datasets aren't prepared for the analysis you want to do. You have to prepare the dataset before you work with it. 

 - We want the average number of schools in a district. 
 - To get that, we need to be able to count the number of schools in each district. 
 - And to do that, we need each school in each district to only appear once. 
 - Last, we only want districts in San Diego county

```{r san diego districts and schools}
sd <- enroll %>% 
  filter(COUNTY == "San Diego") %>% 
  distinct(DISTRICT, SCHOOL) # Each school should only appear once 

View(sd)
```

Question: How many schools does the Alpine district have? 

Count the schools in each district: 

```{r number of schools per district}
school_count <- sd %>% 
  count(DISTRICT, sort = TRUE)

school_count
```


# Pick a Method to Answer the Question 

Compute the average number of schools in a district. Compute the standard deviation of the number of schools in a district. 


# Visualize Data 

Make the dataset easier to visualize by filtering for the top ten: 

```{r top ten}
top_10 <- school_count %>% 
  filter(min_rank(desc(n)) < 10) 

top_10
```

Visualize the top ten districts by number of schools. As I do this, watch how I talk and build the visualization layer by layer based on what I want to see. 

```{r plot number of schools}
ggplot(data = top_10, aes(x = reorder(DISTRICT, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  labs(title = "Districts With the Most Schools", 
       x = "", 
       y = "Number of Schools")
```


# Analyze the Data 

Refresh our memory of the dataset: 

```{r view dataset again}
View(school_count)
```

. . . and let's compute the mean and standard deviation. 

```{r find mean}
school_count %>%
  summarise(avg_schools = mean(n), standard_dev = sd(n), median_schools = median(n))
```


# Report Your Findings

Districts in San Diego have about 16 schools in them, on average. 

We should interpret this with caution, because there's a lot of variation in school counts in San Diego County. San Diego Unified has far and away the highest school count at 223 schools—almost. For the more statistics savvy, that's 7 times the standard deviation—a very large difference from the county average.